Evaluating Deep Learning Methods for Tokenization of Space-less texts in Old French

### 2019-06-04 article/article.tex
Update of the article

\documentclass{jdmdh}
\usepackage[utf8]{inputenc}
\usepackage{array}
\usepackage{pgfplots}
\usepackage{tabularx}
\newcolumntype{C}{>{\arraybackslash}X} % centered "X" column

@@

Tokenization of space-less strings is a task that is specifically difficult for computer when compared to "whathumancando". \textit{Scripta continua} is a writing phenomenon where words would not be separated by spaces and it appears to have disappeared around the 8th century (see \citet{zanna1998lecture}). Never the less, spacing can be somewhat erratic in later centuries writings, as show by Figure \ref{fig:4lines}, a document from the 13th century. In the context of text mining of HTR or OCR output, lemmatization and tokenization of medieval western languages can be a pre-processing step for further research to sustain analyses such as authorship attribution \textbf{CITE JBCAMPS ?}.

@@

\begin{itemize}
    \item 193,734 training examples;
    \item 23,581 validation examples;
    \item 25,512 test examples
\end{itemize}

@@

We believe that, aside from a graphical challenge, word segmentation in OCR from manuscripts can actually be treated from a text point of view and as a NLP task. Word segmentation for some text can be even difficult for humanist, and as such, we believe that post-processing of OCR through tools like Boudams can be a better way to achieve data-mining of the dataset. In light of the high accuracy of the model, we believe the model should perform the same way independently from the language in Medieval Western Europe.

@@

While the baseline surprised us by performing this well on the test corpus, it definitely performed less well than the CNN on a completely unknown corpus: in this context, the proposed model actually shows its ability to carry over unknown corpora in a better way than classical ngram approaches.

@@

\begin{figure}[!ht]
  \centering
  \includegraphics[width=\linewidth]{confusion.png}
  \caption{Confusion matrix of the CNN model without position embedding}
  \label{fig:confusion_matrix}
\end{figure}

>>>>

\documentclass{jdmdh}
\usepackage[utf8]{inputenc}
\usepackage{array}
\usepackage{pgfplots}
\usepackage{float}
\usepackage{tabularx}
\newcolumntype{C}{>{\arraybackslash}X} % centered "X" column

@@

Tokenization of space-less strings is a task that is specifically difficult for computer when compared to "whathumancando". \textit{Scripta continua} is a writing phenomenon where words are be separated by spaces and has disappeared around the 8th century (see \citet{zanna1998lecture}). Never the less, spacing can be somewhat erratic in later centuries writings, as show by Figure \ref{fig:4lines}, a document from the 13th century. In the context of text mining of HTR or OCR output, lemmatization and tokenization of medieval western languages can be a pre-processing step for further research to sustain analyses such as authorship attribution \textbf{CITE JBCAMPS ?}.

@@

\begin{itemize}
    \item 193,734 training examples;
    \item 23,581 validation examples;
    \item 25,512 test examples
    \item Number of classes in testing examples: 482,776 WC; 169,094 WB
    \item Number of classes in unknown examples: 26,393 WC; 10,193 WB
\end{itemize}

@@

We believe that, aside from a graphical challenge, word segmentation in OCR from manuscripts can actually be treated from a text point of view and as a NLP task. Word segmentation for some text can be even difficult for humanist, and as such, we believe that post-processing of OCR through tools like Boudams can be a better way to achieve data-mining of the dataset.

@@

While the baseline surprised us by performing this well on the test corpus, it definitely performed less well than the CNN on a completely unknown corpus: in this context, the proposed model actually shows its ability to carry over unknown corpora in a better way than classical ngram approaches. In light of the high accuracy of the CNN model, we believe the model should perform the same way independently from the language in Medieval Western Europe, and results in annexes on Latin confirms this for at least 2 other corpora.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{confusion.png}
  \caption{Confusion matrix of the CNN model without position embedding}
  \label{fig:confusion_matrix}
\end{figure}

\section{Annex 2 : Scores on Latin Prose and Poetic Corpora}

\begin{table}[H]
\centering
\begin{tabular}{llllllll}
\hline
 & Corpus & Accuracy & Precision & Recall & FScore & WB FN & WB FP \\ \hline
Baseline & Test & 0.978 & 0.961 & 0.974 & 0.968 & 886 & 1893 \\
CNN P & Test & \textbf{0.992} & \textbf{0.987} & \textbf{0.989} & \textbf{0.988} & \textbf{439} & \textbf{584} \\ \hline
Baseline & Unknown & 0.933 & 0.897 & 0.890 & 0.893 & 1587 & 1409 \\
CNN P & Unknown & 0.970 & 0.952 & 0.956 & 0.954 & 600 & 709 \\ \hline
\end{tabular}
\caption{Scores over the Latin classical datasets. FN = False Negative, FP = False Positive}
\label{tab:latin_corpora}
\end{table}

The Latin data is much more noisy than the Old French, as it was less curated than the digital edition we found for Old French. They are part of the Perseus corpus \citet{perseus} and were cut into passages in the context of my thesis. The training, evaluation and test corpora are built upon prose work from Cicero and Suetonius. The unknown corpus is built upon \textit{Epigrammata} from Martial, from book 1 to book 2, as it should be fairly different in word order, vocabulary, etc. Both corpus were generated without noise and word keeping, with a maximum sample size of 150 characters.

\textbf{Statistics:}
\begin{itemize}
\item Number of training examples: 30725
\item Number of evaluation examples: 3558
\item Number of testing examples: 4406
\item Number of classes in testing examples: 105,915 WC; 26,404 WB
\item Number of classes in unknown examples: 35,910 WC; 8,828 WB
\end{itemize}

\section{Annex 3: Scores on Medieval Latin corpora}

\begin{table}[H]
\centering
\begin{tabular}{llllllll}
\hline
 & Corpus & Accuracy & Precision & Recall & FScore & WB FN & WB FP \\ \hline
Baseline & Test & 0.989 & 0.981 & 0.986 & 0.982 & 1036 & 933 \\
CNN P & Test & \textbf{0.997} & \textbf{0.995} & \textbf{0.995} & \textbf{0.995} & \textbf{251} & \textbf{298} \\ \hline
Baseline & Unknown & 0.929 & 0.900 & 0.865 & 0.881 & 14,382 & 27,019 \\
CNN P & Unknown & 0.976 & 0.960 & 0.963 & 0.962 & 6509 & 7444\\ \hline
\end{tabular}
\caption{Scores over the Latin medieval datasets. FN = False Negative, FP = False Positive}
\label{tab:medieval_latin_corpora}
\end{table}

% test : 136,429+1036 WC / 933+33,120 WB
% unk  : 458,273+14,382 WC / 27,019+85,985 WB

The medieval Latin corpora is based on the project Formulae - Litterae - Chartae's open data \citet{formulae} for its training, evaluation and test sets; the unknown corpora is based on three texts from the Monumenta Germanica that are from early to late medieval period (Andreas Agnellus, Manegaldus, Theodoricus de Niem) and are drawn from the Corpus Corporum Project. Both corpus were generated without noise and word keeping, with a maximum sample size of 150 characters. The data presents some MUFI characters but still look like mostly normalized editions, unlike the Old French data.

\textbf{Statistics:}

\begin{itemize}
\item Number of training examples: 36814
\item Number of evaluation examples: 4098
\item Number of testing examples: 5612
\item Number of classes in testing examples: 137,465 WC; 34,053 WB
\item Number of classes in unknown examples: 472,655 WC; 113,004 WB
\end{itemize}
